An Analysis of Blood-Oxygen-Level-Dependent Signal Parameter Estimation Using Particle Filters

### 2011-01-26 doc/isbi/main.tex
updated my ISBI paper

\begin{document}
%\ninept
%
\maketitle
%
\begin{abstract}
The Blood-Oxygen-Level-Dependent (BOLD) signal that is measured by functional 
magnetic resonance imaging (fMRI)
has been the subject of extensive research since the first development 
of the balloon model. While there are definite benefits to 
moving from the Canonical Hemodynamic Response function to a physiologically inspired
BOLD model, significant barriers remain. Optimizing even the simplest
balloon model requires searching within 7 dimensions, and even more complex
models exist. Whereas traditional methods of analyzing fMRI aims
to determine where activation occurs, BOLD models seek a parametric representation
of the signal. Unfortunately, the nonlinear nature of these models
makes it difficult to analyze, therefore this work demonstrates the 
use of a particle filter to regresses the simplest form of the BOLD 
model. The results show that the system of equations are not observable,
leading to a large range of parameters that are consistent with the measurements. 
\end{abstract}
%
\begin{keywords}
BOLD Response, Functional MRI, Nonlinear Systems, Particle Filter, Bayesian Statistics, System Identification
\end{keywords}
%

\begin{acronym}[CMRO2]
\acro{BOLD}{Blood-Oxygen-Level-Dependent}
\acro{CBF}{Cerebral Blood Flow}
\acro{CBV}{Cerebral Blood Volume}
\acro{CMRO2}{Cerebral Metabolic Rate of Oxygen}
\acro{DC}{Direct Current}
\acro{dHb}{Deoxygenated Hemoglobin}
\acro{EM}{Expectation-Maximization}
\acro{EPI}{Echo Planar Imaging}
\acro{fMRI}{Functional Magnetic Resonance Imaging}
\acro{FMRIB}{Oxford Centre for Functional MRI of the Brain}
\acro{FSL}{FMRIB Software Library}
\acro{FWHM}{Full-Width Half-Maximum}
\acro{GA}{Genetic Algorithms}
\acro{GLM}{General Linear Model}
\acro{HRF}{Hemodynamic Response Function}
\acro{Hb}{Hemoglobin}
\acro{MAD}{Median Absolute Deviation}
\acro{MI}{Mutual Information}
\acro{MR}{Magnetic Resonance}
\acro{MRI}{Magnetic Resonance Imaging}
\acro{MSE}{Mean Squared Error}
\acro{O2Hb}{Oxygenated Hemoglobin}
\acro{ODE}{Ordinary Differential Equation}
\acro{PDF}{Probability Density Function}
\acro{POSSUM}{Physics-Oriented Simulated Scanner for Understanding MRI}
\acro{RF}{Radio Frequency}
\acro{RMSE}{Root Mean Squared Error}
\acro{RMSR}{Root Mean Squared Residual}
\acro{SA}{Simulated Annealing}
\acro{SNR}{Signal-to-Noise Ratio}
\acro{SPM}{Statistical Parametric Mapping}
\acro{T1}{Longitudinal}
\acro{T2}{Spin-Spin}
\acro{TR}{Repetition Time}
\acro{UKF}{Unscented Kalman Filter}
\end{acronym}
\section{Introduction}
\label{sec:intro}
Traditional methods of analyzing 
\ac{fMRI} images perform regression using the linear combination of explanatory variables. 
Though moving to a state space model with more more degrees of freedom 
requires additional computation, in this paper Particle Filters will
be used to estimate the governing parameters of the \ac{BOLD} model 
at a computation cost 
that would still allow real time calculations for multiple voxels.
The parameters guiding the BOLD model could provide insight into 
functional differences between patients and brain regions.
More practically, this algorithm is an alternative to the General
Linear Model for activation detection. Though computation intensive,
this method is capable of modeling nonlinear effects
and provides more detailed output.

This work focuses
on the observability of the \ac{BOLD} state equations from Friston et al.
\cite{Friston2000}. 
Because of the nonlinearities in the state equations,
no analytical method exists to determine observability. Therefore
Particle Filters are used to build a joint mixture \ac{PDF} of the model parameters. 
This work will show that this method is sufficient for predicting
the BOLD signal, without requiring excessive computation. 
Because particle filters estimate a full posterior distribution, 
it was possible to calculate correlation of model parameters, an
indicator of observability.

\subsection{BOLD Model}
It is well known that the two types of \ac{Hb} act as contrast agents in 
\ac{EPI} imaging \cite{Buxton1998, WEISSKOFF1994, Ogawa}. Despite this, the connection
between \ac{dHb}/\ac{O2Hb} and neural activity is non-trivial. 
Intuitively, increased 
metabolism will increase \ac{dHb}, however blood vessels are quick
to compensate by permitting increased local blood flow. Increased inflow, accomplished by loosening 
capillary beds, precedes a matching outflow, driving increased 
blood volume.
Since the local \ac{MR} signal depends on the ratio of \ac{dHb} to \ac{O2Hb},
increased blood volume also affects this ratio. 
This was the impetus
for the ground breaking balloon model \cite{Buxton1998} and windkessel
model \cite{Mandeville1999}. These works derive 
the changes in \ac{dHb} ratio and volume of capillaries given \ac{CBF}
and \ac{CMRO2} waveforms.

@@

\section{Methods}
\label{sec:Methods}
\subsection{Particle Filter}
\label{sec:ParticleFilter}
Particle filters, a type of Sequential Monte Carlo (SMC) method,
are a powerful method for on line estimation of the posterior probability 
distribution of parameters given a timeseries and a model. The concept of 
particle filters is similar to that of the Unscented Kalman Filter; however,
distributions are stored as an empirical distribution rather than 
as the first two moments of a Gaussian. Thus, particle filters are 
preferred when the model is nonlinear, and non-Gaussian. 
The particle filters are a well established technique and are 
described in great detail in Arulampalam et al., Thrun et al. 
and Murray \cite{Arulampalam2002a, thrun2008probabilistic, Murray2008}.
\begin{table}[t]
\ninept
  \centering
\begin{tabular}{|c | c  c  c  c  c  c  |}
\hline
  & $\tau_0$ & $\alpha$ & $E_0$    & $V_0$    & $\tau_s$ & $\tau_f$  \\
\hline
$\alpha$                      & 0.880& & & & & \\
\rowcolor[gray]{.8} $E_0$     & -0.766& -0.523& & & & \\
$V_0$                         & 0.624& 0.424& -0.796& & & \\
\rowcolor[gray]{.8} $\tau_s$  & 0.620& 0.295& -0.748& 0.344& & \\
$\tau_f$                      & 0.000& -0.397& -0.431& 0.196& 0.699& \\
\rowcolor[gray]{.8} $\epsilon$& 0.616& 0.656& -0.641& 0.285& 0.446& -0.097\\
\hline
\end{tabular}
  \caption{Correlation Matrix of Posterior Distribution for 40 minute simulated BOLD signal.}
\label{tab:long_corr}
\end{table}
The particle filter was set with each particle representing one possible 
system realization: $\{\tau_0, \alpha, E_0, V_0, \tau_s, \tau_f,
\epsilon, s, f, v, q\}$. Initially 16,000 particles were drawn
from the prior, however after resampling the number was dropped
to 1000. This gave the prior sufficient density to
cover parameter space, but saved computation time once the solution space
became more compact. 
The parameters were drawn independently from the gamma distribution:
$\tau_0 \sim \Gamma(.98, .25)$, 
$\alpha \sim \Gamma(.33, .045)$, $E_0    \sim \Gamma(.34, .03)$,
$V_0    \sim \Gamma(.04, .03)$, $\tau_s \sim \Gamma(1.54, .25)$,
$\tau_f \sim \Gamma(2.46, .25)$, $\epsilon \sim \Gamma(.7, .6)$.
To reduce the size of the search space from 11 to 7, $\{s, f, v, q\}$
were assumed to start at resting state, $\{0, 1, 1, 1\}$. Although
it is common to add noise to the state during integration (to 
emulate noise in the state updates), this was not performed because 
regularization during resampling tended to expand the distribution.
Additionally the cloud of particles is able to account for some amount
of stationary randomness.
Initially each particle is weighted equally, however
as new measurements arrive, they alter the probability
of particles being a solution.
New measurements
are incorporated into the weights by \autoref{eq:weightevolve}:
\begin{equation}
w^i_k & \propto & w^i_{k-1}P(y_k| x^i_k) 
\label{eq:weightevolve}
\end{equation}
where $i$ is the particle index, and $k$ is the time index. $P(y_k | x^i_k)$
is the probability of the observed measurement coming from the $i^{th}$ 
particle and depends on the measurement noise and the observation function, 
\autoref{eq:boldout}.
$x^i_k$ is calculated from $x^i_{k-1}$ using the balloon model, 
\autoref{eq:bold1} through \autoref{eq:bold4}.
For all the analysis  in this work, $1400$ Euler integration points
per TR ($2.1$ s) were used. When the number of particles with significant
(non-zero) weights dropped below 50, regularized resampling was 
applied. Thus, a new set of particles was drawn from the smoothed
version of the mixture \ac{PDF}.

Because fMRI drift can add structure to the noise, a spline with 1 knot every
20 measurements was fitted to and then subtracted from the timeseries
\cite{Smith1999, Tanabe2002}. The level was then converted
to \% difference from the original mean. Since the BOLD signal is 
predominantly positive, a constant (the Gaussian normed \acl{MAD} 
of the \% difference signal) was added to each point.

Given the state-space equations for the \ac{BOLD} signal, simulating a single time
series is straightforward. After generating a true signal,
identically and independently distributed (I.I.D.) Gaussian noise and a Wiener
process with Gaussian I.I.D. steps were added to the true signal. Finally a
carrier level was added, since \ac{BOLD} is 
measured as a \% difference from the baseline. For single voxel
analysis, the noise was kept relatively low ($\sigma = 0.001$  for
I.I.D. Gaussian Noise, and $\sigma = .0005$ for Wiener Steps) to
explore the properties of the model. This noise was added
to the the BOLD signal typically peaks at $0.01-0.03$ (1-3\%). 
To simulate a slice with \ac{POSSUM} \cite{Drobnjak}, it's code was modified to
output activation levels based on sets of parameters rather than
activation levels. For the POSSUM simulated data, SNR was significantly 
lower although the noise did not include drift ($<6dB$).

\section{Results}
\label{sec:Results}
A primary advantage of the particle filter is that it estimates
a complete joint distribution of the parameters and states. This 
allows for more detailed analyses not available to many other methods.
In particular it is possible to calculate the correlation of the
parameters in the posterior distribution. The correlation matrix
for the single 40 minute simulation is shown in \autoref{tab:long_corr}.
\begin{figure}
\begin{minipage}[b]{.5\linewidth}
  \centering
  \centerline{\includegraphics[width=\textwidth]{snr_hm.png}}
  \centerline{(a) SNR}\medskip
\end{minipage}
\hfill
\begin{minipage}[b]{.49\linewidth}
  \centering
  \centerline{\includegraphics[width=\textwidth]{sim_hm_mi.png}}
  \centerline{(b) Mutual Information}\medskip
\end{minipage}
\caption{Results of POSSUM simulated data. $a)$ 
SNR of POSSUM additive noise. $b)$ Mutual Information between
the signal predicted by the particle filter and the actual signal.}
\label{fig:PossumResult}
\end{figure}
\begin{table}[t]
\ninept
  \centering
\begin{tabular}{| c | c | c | c | c | c |}
\hline
& Sim Value & \multicolumn{2}{|c|}{Prior} &\multicolumn{2}{|c|}{Posterior} \\
\hline
        &       &$\mu$  & $\sigma$& $\mu$& $\sigma$\\
\hline
$\tau_0  $& 1.45  & .98   & .25     & 1.07319 & 0.297749  \\
\rowcolor[gray]{.8} 
$\alpha  $&0.3    & .33   & .045    & 0.317955& 0.07311  \\
$E_0     $& 0.47  & .34   & .03     & 0.364195& 0.0457019  \\
\rowcolor[gray]{.8} 
$V_0     $& 0.044 & .04   & .03     & 0.110542& 0.0500039  \\
$\tau_s  $& 1.94  & 1.54  & .25     & 1.86785 & 0.390455  \\
\rowcolor[gray]{.8} 
$\tau_f  $& 1.99  & 2.46  & .25     & 2.50166 & 0.348326  \\
$\epsilon$& 1.8   & .7    & .6      & 1.58716 & 1.14536  \\
\hline
\end{tabular}
\caption{Comparison of the true value (left), the initial (prior) distribution
(center two columns) and the 
mean/variance of the estimates across simulated voxels (Right two Columns).} 
\label{tab:beforeafter}
\end{table}
For POSSUM simulations, \autoref{fig:PossumResult} shows the
the mutual information between the estimated timeseries and the 
preprocessed signal. Note that in order to remove all false
positives a threshold of $0.15$ bits was applied, 
the regions of significant SNR are visible despite this.
Each estimate was generated from the 
mean of the particles' parameters at the end of the run. 
Additionally the mean and variance of the parameter estimates
(mean) across the spatial regions of \autoref{fig:PossumResult} are
given in \autoref{tab:beforeafter}.

\section{Discussion}
\label{sec:Conclusion}
The data shows substantial correlation present in the parameter 
estimates for the 40 minute simulated data. Correlation of 
constants is a definite sign of the model being under-determined
and thus not observable. This is in spite of low noise, and the 
particle filter's solution converging to a Root-Mean-Squared error
of less than $0.002$.

On a larger scale, the particle filter also performed well, clearly
locating regions of activation in spite of Signal-to-Noise Ratio 
well below 6dB. However, the final parameter estimates 
varied more than the initial (prior) distribution (\autoref{tab:beforeafter}).
This effect is slightly deceiving because the distributions
were far from Gaussian, and many were bimodal. It is 
significant, however, that such a wide range of solutions
could give good estimates for the true BOLD signal.

From the results, it is clear that the BOLD model as presented
by Friston et al. is not observable, and thus parameter estimates
suffer from very large variance \cite{Friston2000}. Therefore
any work attempting to estimate these parameters from the BOLD
signal alone is likely measuring the variance due the model
rather than the variance due to try physiological differences. 
There are two ways to deal with this problem. First, as
specified in Deneux et al., certain parameters could be held constant
\cite{Deneux2006}. This solution is relatively easy, and would
also reduce computation time. A better way of dealing with this 
problem is simultaneously measuring fMRI and \ac{CBF}.  
This additional measurement, which is closer to 
original stimulus, could make the Balloon Model observable.

>>>>

\begin{document}
%\ninept
%
\maketitle
%
\begin{abstract}
The Blood-Oxygen-Level-Dependent (BOLD) signal that is measured by functional 
magnetic resonance imaging (fMRI)
has been the subject of extensive research since the development 
of the first balloon model. While there are definite benefits to 
moving from the Canonical Hemodynamic Response function to a physiologically inspired
BOLD model, significant barriers remain. Optimizing the simplest
balloon model requires searching within 7 dimensions, and even more complex
models exist. Additionally, the nonlinear nature of these models
make them difficult to analyze, therefore this work 
uses a particle filter to regresses a simple form of the BOLD 
model. Whereas traditional methods of analyzing fMRI aims
to determine where activation occurs, BOLD model regression seeks a parametric representation
of the signal. The results show that the particle filter 
attains a good fit but that the system of equations are not observable,
leading to a large range of parameters that are consistent with the measurements. 
\end{abstract}
%
\begin{keywords}
BOLD Response, Functional MRI, Nonlinear Systems, Particle Filter, Bayesian Statistics, System Identification
\end{keywords}
%

\begin{acronym}[CMRO2]
\acro{BOLD}{Blood-Oxygen-Level-Dependent}
\acro{CBF}{Cerebral Blood Flow}
\acro{CBV}{Cerebral Blood Volume}
\acro{CMRO2}{Cerebral Metabolic Rate of Oxygen}
\acro{DC}{Direct Current}
\acro{dHb}{Deoxygenated Hemoglobin}
\acro{EM}{Expectation-Maximization}
\acro{EPI}{Echo Planar Imaging}
\acro{fMRI}{Functional Magnetic Resonance Imaging}
\acro{FMRIB}{Oxford Centre for Functional MRI of the Brain}
\acro{FSL}{FMRIB Software Library}
\acro{FWHM}{Full-Width Half-Maximum}
\acro{GA}{Genetic Algorithms}
\acro{GLM}{General Linear Model}
\acro{HRF}{Hemodynamic Response Function}
\acro{Hb}{Hemoglobin}
\acro{MAD}{Median Absolute Deviation}
\acro{MI}{Mutual Information}
\acro{MR}{Magnetic Resonance}
\acro{MRI}{Magnetic Resonance Imaging}
\acro{MSE}{Mean Squared Error}
\acro{O2Hb}{Oxygenated Hemoglobin}
\acro{ODE}{Ordinary Differential Equation}
\acro{PDF}{Probability Density Function}
\acro{POSSUM}{Physics-Oriented Simulated Scanner for Understanding MRI}
\acro{RF}{Radio Frequency}
\acro{RMSE}{Root Mean Squared Error}
\acro{RMSR}{Root Mean Squared Residual}
\acro{SA}{Simulated Annealing}
\acro{SNR}{Signal-to-Noise Ratio}
\acro{SPM}{Statistical Parametric Mapping}
\acro{T1}{Longitudinal}
\acro{T2}{Spin-Spin}
\acro{TR}{Repetition Time}
\acro{UKF}{Unscented Kalman Filter}
\end{acronym}
\section{Introduction}
\label{sec:intro}
Traditional methods of analyzing 
\ac{fMRI} images perform regression using the linear combination of explanatory variables. 
Moving to a state-space model with more degrees of freedom 
requires additional computation; however, in this paper Particle Filters will
be used to estimate the governing parameters of the \ac{BOLD} model 
at a computation cost 
that would still permit real time calculations for multiple voxels.
The parameters guiding the BOLD model may provide insight into 
functional differences between patients and brain regions.
More practically, this algorithm provides a non-parametric
alternative to the General Linear Model. Though computationally intensive,
this method is capable of modeling nonlinear effects
and provides more detailed output.

This work focuses
on the observability of the \ac{BOLD} state equations from Friston et al.
\cite{Friston2000}. 
Because of the nonlinearities in the state equations,
no analytical method exists to determine observability. Therefore
Particle Filters are used to build a joint mixture \ac{PDF} of the model parameters. 
This work will show that the method is sufficient for predicting
the BOLD signal, without requiring excessive computation. 
Because particle filters estimate a full posterior distribution, 
it was possible to calculate correlation of model parameters, an
indicator of observability.

\subsection{BOLD Model}
It is well known that the two types of \ac{Hb} act as contrast agents in 
\ac{EPI} imaging \cite{Buxton1998, WEISSKOFF1994, Ogawa}. Despite this, the connection
between \ac{dHb}/\ac{O2Hb} and neural activity is complex. 
Intuitively, increased 
metabolism increases \ac{dHb}, however blood vessels quickly
compensate by permitting increased local blood flow. Increased inflow, 
precedes a matching outflow, which causes increased blood volume.
Since the local \ac{MR} signal depends on the ratio of \ac{dHb} to \ac{O2Hb},
increased blood volume also affects this ratio. 
This relation of \ac{dHb}, local blood flow and local blood volume,
was the impetus
for the ground breaking balloon model \cite{Buxton1998} and windkessel
model \cite{Mandeville1999}. These works derive 
the changes in \ac{dHb}/\ac{O2Hb} and blood volume given waveforms for \ac{CBF}
and \ac{CMRO2}.

@@

\section{Methods}
\label{sec:Methods}
\subsection{Particle Filter}
\label{sec:ParticleFilter}
Particle filters, a type of Sequential Monte Carlo (SMC) method,
are a powerful method for on line estimation of the posterior probability 
distribution of parameters given a timeseries and a model. The concept of 
particle filters is similar to that of the Unscented Kalman Filter; however,
distributions are stored as an empirical distribution rather than 
as the first two moments of a Gaussian. Thus, particle filters are 
preferred when the model is non-Gaussian and nonlinear. 
The particle filters are a well established technique and are 
described in great detail in Arulampalam et al., Thrun et al. 
and Murray \cite{Arulampalam2002a, thrun2008probabilistic, Murray2008}.
\begin{table}[t]
\ninept
  \centering
\begin{tabular}{|c | c  c  c  c  c  c  |}
\hline
  & $\tau_0$ & $\alpha$ & $E_0$    & $V_0$    & $\tau_s$ & $\tau_f$  \\
\hline
$\alpha$                      & 0.880& & & & & \\
\rowcolor[gray]{.8} $E_0$     & -0.766& -0.523& & & & \\
$V_0$                         & 0.624& 0.424& -0.796& & & \\
\rowcolor[gray]{.8} $\tau_s$  & 0.620& 0.295& -0.748& 0.344& & \\
$\tau_f$                      & 0.000& -0.397& -0.431& 0.196& 0.699& \\
\rowcolor[gray]{.8} $\epsilon$& 0.616& 0.656& -0.641& 0.285& 0.446& -0.097\\
\hline
\end{tabular}
  \caption{Correlation Matrix of Posterior Distribution for 40 minute simulated BOLD signal.}
\label{tab:long_corr}
\end{table}
The particle filter was set with each particle representing one possible 
system realization: $\{\tau_0, \alpha, E_0, V_0, \tau_s, \tau_f,
\epsilon, s, f, v, q\}$. Initially 16,000 particles were drawn
from the prior, however after resampling the number was dropped
to 1000. This gave the prior sufficient density to
cover parameter space, but saved computation time once the solution space
became more compact. 
The parameters were drawn independently from the gamma distribution:
$\tau_0 \sim \Gamma(.98, .25)$, 
$\alpha \sim \Gamma(.33, .045)$, $E_0    \sim \Gamma(.34, .03)$,
$V_0    \sim \Gamma(.04, .03)$, $\tau_s \sim \Gamma(1.54, .25)$,
$\tau_f \sim \Gamma(2.46, .25)$, $\epsilon \sim \Gamma(.7, .6)$.
To reduce the size of the search space from 11 to 7, $\{s, f, v, q\}$
were assumed to start at resting state, $\{0, 1, 1, 1\}$. Although
it is common to add noise to the state during integration (to 
emulate noise in the state updates), this was not performed because 
a) regularization during resampling tended to expand the distribution
and b) the cloud of particles is able to account for some amount
of stationary randomness.
Initially each particle is weighted equally, however
as new measurements arrive, they alter the probability
of particles being a solution.
New measurements
are incorporated into the weights by \autoref{eq:weightevolve}:
\begin{equation}
w^i_k & \propto & w^i_{k-1}P(y_k| x^i_k) 
\label{eq:weightevolve}
\end{equation}
where $i$ is the particle index, and $k$ is the time index. $P(y_k | x^i_k)$
is the probability of the observed measurement coming from the $i^{th}$ 
particle and depends on the measurement noise and the observation function, 
\autoref{eq:boldout}.
$x^i_k$ is calculated from $x^i_{k-1}$ using the balloon model, 
\autoref{eq:bold1} through \autoref{eq:bold4}.
For all the analysis  in this work, $1400$ Euler integration points
per TR ($2.1$ s) were used. When the number of particles with significant
(non-zero) weights dropped below 50, regularized resampling was 
applied. Thus, a new set of particles was drawn from the smoothed
version of the mixture \ac{PDF}.

Because fMRI drift can add structure to the noise, a spline fit with 1 knot every
40 seconds was subtracted from the timeseries
\cite{Smith1999, Tanabe2002}. Since the BOLD signal is 
predominantly positive, a constant (the Gaussian-normed \acl{MAD})
was added to each point.

Given the state-space equations for the \ac{BOLD} signal, simulating a single time
series is straightforward. After generating a true signal,
identically and independently distributed (I.I.D.) Gaussian noise and a Wiener
process with Gaussian I.I.D. steps were added to the true signal. Finally a
carrier level was added, since \ac{BOLD} is 
measured as a \% difference from the baseline. For single voxel
analysis, the noise was kept relatively low ($\sigma = 0.001$  for
I.I.D. Gaussian Noise, and $\sigma = .0005$ for Wiener Steps) to
explore the properties of the model. This noise was added
to the BOLD signal which has a typical RMS of $0.01-0.03$ (1-3\%). 
To simulate a slice with \ac{POSSUM} \cite{Drobnjak}, it's code was modified to
output activation levels based on sets of parameters rather than
activation levels. For the POSSUM simulated data, SNR was significantly 
lower although the noise did not include drift ($<6dB$).

\section{Results}
\label{sec:Results}
A primary advantage of the particle filter is that it estimates
a complete joint distribution of the parameters and states. This 
allows for more detailed analyses not available to many other methods.
In particular it is possible to calculate the correlation of the
parameters in the posterior distribution. The correlation matrix
for the single 40 minute simulation is shown in \autoref{tab:long_corr}.
\begin{figure}
\begin{minipage}[b]{.5\linewidth}
  \centering
  \centerline{\includegraphics[width=\textwidth]{snr_hm.png}}
  \centerline{(a) SNR}\medskip
\end{minipage}
\hfill
\begin{minipage}[b]{.49\linewidth}
  \centering
  \centerline{\includegraphics[width=\textwidth]{sim_hm_mi.png}}
  \centerline{(b) Mutual Information}\medskip
\end{minipage}
\caption{Results of POSSUM simulated data. $a)$ 
SNR of POSSUM additive noise. $b)$ Mutual Information between
the signal predicted by the particle filter and the actual signal.}
\label{fig:PossumResult}
\end{figure}
\begin{table}[t]
\ninept
  \centering
\begin{tabular}{| c | c | c | c | c | c |}
\hline
& Sim Value & \multicolumn{2}{|c|}{Prior} &\multicolumn{2}{|c|}{Posterior} \\
\hline
        &       &$\mu$  & $\sigma$& $\mu$& $\sigma$\\
\hline
$\tau_0  $& 1.45  & .98   & .25     & 1.07319 & 0.297749  \\
\rowcolor[gray]{.8} 
$\alpha  $&0.3    & .33   & .045    & 0.317955& 0.07311  \\
$E_0     $& 0.47  & .34   & .03     & 0.364195& 0.0457019  \\
\rowcolor[gray]{.8} 
$V_0     $& 0.044 & .04   & .03     & 0.110542& 0.0500039  \\
$\tau_s  $& 1.94  & 1.54  & .25     & 1.86785 & 0.390455  \\
\rowcolor[gray]{.8} 
$\tau_f  $& 1.99  & 2.46  & .25     & 2.50166 & 0.348326  \\
$\epsilon$& 1.8   & .7    & .6      & 1.58716 & 1.14536  \\
\hline
\end{tabular}
\caption{Comparison of the true value (left), the initial (prior) distribution
(center two columns) and the 
mean/variance of the estimates across simulated voxels (Right two Columns).} 
\label{tab:beforeafter}
\end{table}
For POSSUM simulations, \autoref{fig:PossumResult} shows the
the mutual information between the estimated timeseries and the 
preprocessed signal. Note that in order to remove all false
positives a threshold of $0.15$ bits was applied.
Each estimate was generated from the 
mean of the particles' parameters at the end of the run. 
Additionally the mean and variance of the parameter estimates
(mean) across the spatial regions of \autoref{fig:PossumResult} are
given in \autoref{tab:beforeafter}.

\section{Discussion}
\label{sec:Conclusion}
The single voxel simulation data (\autoref{tab:long_corr}) 
shows substantial correlation present in the parameter 
estimates for the 40 minute simulated data. Correlation of 
constants is a definite sign of the model being under-determined
and thus not observable. This is in spite of low noise, and the 
particle filter's solution (mean) converging to a Root-Mean-Squared error
of less than $0.002$.

In the \ac{POSSUM} simulation, the particle filter clearly
located regions of activation in spite of a Signal-to-Noise Ratio 
well below 6dB. However, the variance of the final parameter estimates 
was greater than the initial (prior) distribution (\autoref{tab:beforeafter}).
This effect is slightly deceiving because the distributions
were far from Gaussian, and many were bimodal. It is 
important, however, that such a wide range of solutions
could give good estimates for the true BOLD signal.

From single voxel and single slice results, 
there is substantial evidence that the BOLD model as presented
by Friston et al. is not observable, and thus parameter estimates
suffer from very large variance \cite{Friston2000}.  
There are two ways to deal with this problem. First, as
specified in Deneux et al., certain parameters could be held constant
\cite{Deneux2006}. This solution is relatively easy, and would
also reduce computation time. A better way of dealing with this 
problem is simultaneously measuring fMRI and \ac{CBF}.  
This additional measurement, which is closer to 
original stimulus, could make the Balloon Model observable.
